Strong lensing selection effects

### 2022-07-31 paper/seleff.tex
mocks
experimenting with different lens definitions

@@

\begin{figure*}
\includegraphics[width=\textwidth]{footprints.eps}
\caption{
Criterion used to classify lensed images of extended sources. Four examples. In each panel, the blue regions are the footprint of the lensed source. 
The two segments connecting the source footprint to the lens centre show the largest angle subtended by the source. If this angle is larger than $90^\circ$, the system is classified as a strong lens.
\label{fig:lensdef}
}
\end{figure*}
%__________________________________________________________________

@@

%After comparing the predicted and observed fundamental plane scatter for different sets of values of $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$, we settled on three 
%After comparing the predicted and observed fundmantal plane scatter, 
We settled on three different set of intrinsic scatter parameters, as indicated in \Tref{tab:scatter}. We label them the fiducial, the low-scatter and the high-scatter scenarios.
The high-scatter scenario is ruled out both by the fundamental plane and by weak lensing constraints \citep{Son++22}. 
Moreover, our dynamical model is quite simplistic, as it neglects the effects of orbital anisotropy and departures from spherical symmetry, which are additional sources of scatter. 
Nevertheless, we use it in our experiment in order to obtain a more conservative upper limit on the amplitude of the strong lensing bias.
%A model with the same intrinsic scatter parameters and a more realistic 3-dimensional and orbital structure would likely predict a fundamental plane scatter that is at least as large as the observed one.
\begin{table}
\caption{Intrinsic scatter scenarios}
\label{tab:scatter}
\begin{tabular}{lccc}
\hline
\hline
Model name & $\sigma_{\mathrm{sps}}$ & $\sigma_{\mathrm{h}}$ & Predicted FP \\
 & & & scatter \\
\hline
Fiducial & 0.08 & 0.20 & 0.034 \\
Low scatter & 0.05 & 0.10 & 0.021 \\
High scatter & 0.10 & 0.30 & 0.433 \\
\end{tabular}
\notes{
Adopted values of the intrinsic scatter parameters $\sigma_{\mathrm{sps}}$ and $\sigma_{\mathrm{h}}$ in different simulations. For each set of values, the fourth column indicates the scatter around the fundamental plane predicted via the spherical Jeans analysis of Appendix~\ref{sect:appendixb}.
The observed fundamental plane scatter is $0.035$.
}
\end{table}

@@

\ale{To be filled by Shun-Sheng}.
We applied a lower limit to the source redshift, by selecting only sources with $z_s > 0.8$.
This ensures that all of the sources lie behind all of the lenses.
A similar cut could be applied to a real survey using photometric redshift, to reduce the incidence of false positives (e.g. arc-like features physically associated with the lens galaxy) in the lens finding phase.
%This selection is meant to simulate a cut on the redshift (possibly photometric) of the arcs, which would help reduce the incidence of false positives during the lens finding phase.
The resulting number density of sources is ZZZ~arcmin$^{-2}$.
We approximated their spatial distribution as uniform in the sky, that is we neglected clustering of the sources.

>>>>

\def\msource{m_{\mathrm{s}}}

@@

%\begin{figure*}
%\includegraphics[width=\textwidth]{footprints.eps}
%\caption{
%Criterion used to classify lensed images of extended sources. Four examples. In each panel, the blue regions are the footprint of the lensed source. 
%The two segments connecting the source footprint to the lens centre show the largest angle subtended by the source. If this angle is larger than $90^\circ$, the system is classified as a strong lens.
%\label{fig:lensdef}
%}
%\end{figure*}
\begin{figure}
\includegraphics[width=\columnwidth]{example_detections.eps}
\caption{
Criterion used to classify lensed images of extended sources. Four examples.
\label{fig:lensdef}
}
\end{figure}

@@

%After comparing the predicted and observed fundamental plane scatter for different sets of values of $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$, we settled on three 
%After comparing the predicted and observed fundmantal plane scatter, 
We settled on three different set of intrinsic scatter parameters, as indicated in \Tref{tab:scatter}. We label them the fiducial, the low-scatter and the high-scatter scenarios.
The high-scatter scenario is ruled out both by the fundamental plane and by weak lensing constraints \citep{Son++22}. 
Moreover, our dynamical model is quite simplistic, as it neglects the effects of orbital anisotropy and departures from spherical symmetry, which are additional sources of scatter. 
Nevertheless, we use it in our experiment in order to obtain a more conservative upper limit on the amplitude of the strong lensing bias.
%A model with the same intrinsic scatter parameters and a more realistic 3-dimensional and orbital structure would likely predict a fundamental plane scatter that is at least as large as the observed one.
\begin{table}
\caption{Intrinsic scatter scenarios}
\label{tab:scatter}
\begin{tabular}{lccc}
\hline
\hline
Model name & $\sigma_{\mathrm{sps}}$ & $\sigma_{\mathrm{h}}$ & Predicted FP \\
 & & & scatter \\
\hline
Fiducial & 0.08 & 0.20 & 0.034 \\
Low scatter & 0.05 & 0.10 & 0.021 \\
High scatter & 0.10 & 0.30 & 0.433 \\
\end{tabular}
\tablefoot{
Adopted values of the intrinsic scatter parameters $\sigma_{\mathrm{sps}}$ and $\sigma_{\mathrm{h}}$ in different simulations. For each set of values, the fourth column indicates the scatter around the fundamental plane predicted via the spherical Jeans analysis of Appendix~\ref{sect:appendixb}.
The observed fundamental plane scatter is $0.035$.
}
\end{table}

@@

%\ale{To be filled by Shun-Sheng}.
We applied a lower limit to the source redshift, by selecting only sources with $z_s > 0.8$.
This ensures that all of the sources lie behind all of the lenses.
A similar cut could be applied to a real survey using photometric redshift, to reduce the incidence of false positives (e.g. arc-like features physically associated with the lens galaxy) in the lens finding phase.
%This selection is meant to simulate a cut on the redshift (possibly photometric) of the arcs, which would help reduce the incidence of false positives during the lens finding phase.
The resulting number density of sources is $70$~arcmin$^{-2}$.
We approximated their spatial distribution as uniform in the sky, that is we neglected clustering of the sources.
